Secure Composition \\ of Robust and Optimising Compilers

### 2024-12-10 icfp/main.tex
Add icfp folder

% Fixing: Too many math alphabets used in version normal
\newcommand\hmmax{0}
\newcommand\bmmax{0}
\documentclass[dvipsnames,conference]{IEEEtran}

@@

\usepackage[colorinlistoftodos]{todonotes}
\usepackage[inference]{semantic}
\usepackage[switch]{lineno}
\usepackage{halloweenmath}
\usepackage{fontawesome5}
\usepackage{listofitems}
\usepackage{breakcites}
\usepackage{glossaries}
\usepackage{hyperref}
\usepackage{cleveref}
\usepackage{stmaryrd}
\usepackage{marvosym}
\usepackage{listings}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{nameref}
\usepackage{amsthm}
\usepackage{xspace}
\usepackage{xfrac}
\usepackage{tikz}
\usepackage{soul}
\usepackage{bm}
\usepackage{paralist}

@@

\loadglsentries{acronyms}
\makeglossaries

\AtBeginDocument{%
  \providecommand\BibTeX{{%
    \normalfont B\kern-0.5em{\scshape i\kern-0.25em b}\kern-0.8em\TeX}}}
\makeatletter
\newcommand{\linebreakand}{%
  \end{@IEEEauthorhalign}
  \hfill\mbox{}\par
  \mbox{}\hfill\begin{@IEEEauthorhalign}
}
\makeatother

\renewcommand\thelinenumber{\color{red}\arabic{linenumber}}
\begin{document}
\linenumbers

\title{
  Secure Composition 
  \\ 
  of Robust and Optimising Compilers
  \vspace{-1.5em}
}

% \author{
% Matthis Kruse\orcidID{0000-0003-4062-9666}
% \and
% Michael Backes
% \and
% Marco Patrignani
% }
% \authorrunning{M. Kruse et al.}

@@

% \begin{IEEEkeywords}
%   memory-safety, cryptographic constant-time, secure compilation, robustness
% \end{IEEEkeywords}

@@

  \item %
  		This paper provides formal proofs of the security properties preserved by each of the presented passes. 
  		% 
  		Additionally this paper analyses each pass in terms of their compatibility with each other (\Cref{sec:formalities}), proving how does each pass fulfill the conditions that make their composition meaningful from a security standpoint.
        % In doing so, the paper provides formal insights into some technicalities of the concrete proofs necessary as technical setup to reap the benefits of the presented compositionality framework.

@@

Properties $\pi$ are sets of traces of admissible program behaviours, ascribing what said property considers valid.
% 
The set of all properties can be partitioned into different {\em classes} ($\class$), i.e., safety, liveness, and neither safety nor liveness~\cite{clarkson2008hyper}, so a class is simply a set of properties.
% 
The compositionality framework (\Cref{sec:sequential}) presented in this paper considers arbitrary classes, while the case-study (\Cref{sec:casestud:rtp}) fixes them to concrete instances of safety properties, since it is decidable whether a trace satisfies a safety property with just a finite trace (i.e., a \emph{prefix}).
As an example, consider the trace:
$$\ev{Dealloc\ \loc}\cdot\ev{Read\ \loc\ 1729}\cdot\dots$$
which describes the interaction with a memory where the deallocation of an address $\loc$ precedes a read (of some value $1729$) at that address in memory.
% 
This program behaviour is insecure w.r.t. a canonical notion of (temporal) memory safety dictating no use-after-frees of pointers~\cite{nagarakatte2010cets,azevedo2018meaningsofms}, because it reads from a memory location that was freed already.
The previous finite trace prefix is enough to decide that the trace does not satisfy \gls*{tms} and there is no way to append events to this prefix which would result in the trace being admissible.

@@

This paper uses a general notion of robust preservation~\cite{abate2021extacc} that considers compilers that use languages with different trace models. 
To this end, considering a source trace $\src{\trace}$ and a target trace $\trg{\trace}$, there is a relation ($\sim$) describing the effect of a corresponding compiler (see \Cref{sec:casestud:rtp}). 
This relation induces the following two projection functions~\cite{abate2021extacc}: (1) the \emph{existential image} $\tau_\sim\left(\src{\pi}\right)$ and (2) the \emph{universal image} $\sigma_\sim\left(\trg{\pi}\right)$.
These projections map source-level (resp. target-level) properties to target-level (resp. source-level) properties in a way that identifies the ``same'' property across languages. 
The case study of this paper uses the universal image, since some considered properties, such as \gls*{ss}, are not definable in a higher-level language that, e.g., does not model speculation.
% For self-containment, the universal image is defined as follows~\cite{abate2021extacc}:
% 
\begin{definition}[Universal Image]
% [Existential and Universal Image]
\label{def:universal:img}\label{def:sigma}
  % \[
  %   \tau_\sim\left(\src{\pi}\right) := 
  %     \left\{ 
  %       \trg{\trace} \mid \exists \src{\trace}\ldotp \src{\trace}\sim\trg{\trace}, \text{ and } \src{\trace}\in\src{\pi} 
  %     \right\}
  % \]
  \[ 
    \sigma_\sim\left(\trg{\pi}\right) := 
      \left\{ 
        \src{\trace} \mid \forall \trg{\trace}\ldotp \text{if }\src{\trace}\sim\trg{\trace}, \text{ then } \trg{\trace}\in\trg{\pi} 
      \right\}
  \]
\end{definition}
\noindent
With this projection function, we define a more general version of robust preservation as follows~\cite{abate2021extacc}.
% For a \bul{compiler $\cc{\src{L}}{\trg{L}}$ to robustly preserve a class of source properties $\src{\class}$}, if for any \rul{property $\src{\pi}$ of that class $\src{\class}$ and programs $\src{p}$ written in $\src{L}$} where \iul{$\src{\progvar}$ robustly satisfies $\src{\pi}$}, then \oul{the compilation of $\src{\progvar}$, $\cc{\src{L}}{\trg{L}}\left(\src{p}\right)$, must robustly satisfy $\tau_\sim\left(\src{\pi}\right)$}.
% 
% \begin{definition}[Robust Preservation with $\tau_\sim$]\label{def:rtp:tau}
%   % $\;$\\
%   % \vspace{-1em}
%   % \begin{nscenter}
%   % Compiler $\cc{\src{L}}{\trg{L}}$ robustly preserves $\class$, 
%   \bul{$\rtptau{\cc{\src{L}}{\trg{L}}}{\src{\class}}{\sim}$}
%   %, iff 
%   $\isdef$
%   \rul{$\forall \src{\pi}\in\src{\class}, \src{p}\in\src{L},$} if \iul{$\rsat{\src{\progvar}}{\src{\pi}}$}, then \oul{$\rsat{\cc{\src{L}}{\trg{L}}\left(\src{p}\right)}{\tau_\sim{\left(\src{\pi}\right)}}$}.
%   % \end{nscenter}
% \end{definition}
% 
A \bul{compiler $\cc{\src{L}}{\trg{L}}$ robustly preserves a class of target properties $\trg{\class}$}, if for any \rul{property $\trg{\pi}$ of class $\trg{\class}$ and programs $\src{p}$}, where \iul{$\src{\progvar}$ robustly satisfies $\sigma_\sim\left(\trg{\pi}\right)$}, \oul{the compilation of $\src{\progvar}$, we have that $\cc{\src{L}}{\trg{L}}\left(\src{p}\right)$ robustly satisfies $\trg{\pi}$}.

@@

Let \bul{$\cc{\src{L}}{\trg{L}}$ robustly preserve the class $\sigma_{\sim_2}\left(\trg{\class[_1]}\right)$ under $\sim_1$} and let \rul{$\cc{\trg{L}}{\obj{L}}$ robustly preserve the class $\trg{\class[_2]}$ under $\sim_2$}.
Then, \iul{when the cross-language relation $\sim_2$ is well-formed w.r.t. class $\trg{\class[_1]}$}, it follows that \oul{the composed compiler $\cc{\src{L}}{\trg{L}}\circ\cc{\trg{L}}{\obj{L}}$ robustly preserves the intersection of classes $\trg{\class[_1]}\cap\trg{\class[_2]}$ under $\sim_1\bullet\sim_2$}.
% 
%For space reasons, \Thmref{thm:urtp} and \Thmref{thm:lrtp} are not presented here, but can be found in the technical appendix ($\CoqSymbol$).
% Note that it is not necessary to require $\sim_2$ to be well-formed w.r.t. $\src{\class[_1]}$, since the 
% \MPin{why}
\begin{theorem}[Composition of Secure Compilers w.r.t. $\sigma$]\label{thm:rtpsim:sig}
  $\;$

  If \bul{$\rtpsig{\cc{\src{L}}{\trg{L}}}{\tilde{\sigma}_{\sim_2}\left(\trg{\class[_{1}]}\right)}{\sim_1}$}, \rul{$\rtpsig{\cc{\trg{L}}{\obj{L}}}{\trg{\class[_2]}}{\sim_2}$}, and \iul{$\wfcsig{\sim_2}{\trg{\class[_1]}}$}, \\ then \oul{$\rtpsig{\cc{\src{L}}{\trg{L}}\circ\cc{\trg{L}}{\obj{L}}}{\trg{\class[_{1}]}\cap\trg{\class[_{2}]}}{\sim_1\bullet\sim_2}$}. \Coqed
\end{theorem}

@@

This section introduces trace properties of interest for this paper: \gls*{tms}, \gls*{sms}, \gls*{ms}, \gls*{scct}, and \gls*{ss}.
These properties are of practical importance (as mentioned in \Cref{sec:introduction}) and also of interest in the case study presented later (\Cref{sec:casestud:defs,sec:casestud:rtp}). 
This section presents all of them, despite the fact that they are inspired by existing work, in order to showcase all that is required for a formal proof of security for a realistic compilation toolchain.
%\MP{ say something about this in the rebuttal: only the defs are borrowed, the universal images and all that work is new}
%\MK{ universal images are not new. }
% , none of the properties is novel, for self-containment of the paper, their definition is necessary for the case study.
% For each of the presented properties, the technical report defines corresponding monitors, which refine a given property and have a key role in the proofs of this paper. 
The technical report defines monitors for each of the presented properties.
% 
Monitors refine each property and have a key tole in the proofs of this paper.
% managed to prove 
%  if tms-monitor --[trace]-->* ∅  
%  then trace ∈ tms
% but not
%  if trace ∈ tms
%  then tms-monitor --[trace]-->* ∅

@@

To express Constant Time, we extend the memory safety trace model with a {\em security tag} ($\securitytag{}$) that indicates whether events contains sensitive information ($\lock$) or not ($\unlock$).

@@

\[
  \typerulenolabel{specrel:drop:tag}{}{\preevent[_{\ctsafe}];\lock\sim_{\mathghost}\preevent[_{\ctsafe}];\lock_{\text{NONE}}}
  \typerulenolabel{specrel:drop:spectag}{}{\emptyevent\sim_{\mathghost}\preevent[_{\mathghost}];\lock_{\text{PHT}}}
\]
\[
  \typerulenolabel{specrel:drop:pub}{}{\preevent[_{\ctsafe}];\unlock\sim_{\mathghost}\preevent[_{\ctsafe}];\unlock}
  \typerulenolabel{specrel:drop:crash}{}{\lightning\sim_{\mathghost}\lightning}
\]
\[
  \typerulenolabel{specrel:drop:spec}{}{\emptyevent\sim_{\mathghost}\ev{Spec}}
  \typerulenolabel{specrel:drop:rlb}{}{\emptyevent\sim_{\mathghost}\ev{Rlb}}
  \typerulenolabel{specrel:drop:barrier}{}{\emptyevent\sim_{\mathghost}\ev{Barrier}}
\]

@@

% \vspace{-1em}
\[
  \begin{array}{rrcl}
    \text{(Values)} & \trg{v} &:=& \dots \mid \trg{\lpair{v_1}{v_2}} \\
    \text{(Expressions)} & \trg{\expr} &:=& \dots \mid \trg{\lhast{\expr}{\tau}} \mid \trg{\lispoisoned{x}}
  \end{array}
\]

@@

    \draw[dashed] ($(S)-(0.0,-0.5)$) -- ($(S)-(0.0,-2.5)$);
    \draw[dashed] ($(T)-(0.0,-0.5)$) -- ($(T)-(0.0,-2.5)$);
    \draw[dashed] ($(M)-(0.0,-0.5)$) -- ($(M)-(0.0,-2.5)$);
    \draw[dashed] ($(E)-(0.0,-0.5)$) -- ($(E)-(0.0,-2.5)$);
    \draw[dashed] ($(O)-(0.0,-0.5)$) -- ($(O)-(0.0,-2.5)$);
    \draw[dashed] ($(Z)-(0.0,-0.5)$) -- ($(Z)-(0.0,-2.5)$);
  \end{tikzpicture}
  \vspace{-1em}
  \caption{Visualisation of the optimising compilation pipeline that preserves \gls*{specms}. %
    Vertices in the graph are the programming languages from earlier sections (\Cref{sec:casestud:defs}). %
    Full edges are secure compilers passes.
    Dotted edges are composition of passes and use the presented framework (\Cref{sec:sequential}) to indicate the property they preserve. %
    The dashed lines partition the graph into the sections where the respective theorems are presented.
  }\label{fig:pipeline}
\end{figure*}
The section demonstrates the power of the framework (\Cref{sec:sequential}) by composing these compilers for a secure and optimising compilation chain that robustly preserves \gls*{specms}.
The first step in this chain is the compiler from $\src{L_{\tmssafe}}$ to $\trg{L}$ that robustly preserves just \gls*{tms} (\Cref{thm:cca:rtp:tms}).
From here, another pass from $\trg{L}$ to $\irl{L_{\mssafe}}$ ensures that no out-of-bounds accesses can happen and, thus, programs at this point attain \gls*{sms} (\Cref{thm:ccb:rtp:sms}).
Since these properties compose into \gls*{ms}, composing these passes yields a compiler that robustly preserves \gls*{ms} (\Cref{corr:ccab:rtp:ms}).
Then, the section presents two optimisation passes, namely \gls*{cf} and \gls*{dce}, each of which robustly preserves \gls*{ms} (\Cref{thm:ccdce:rtp:ms,thm:cccf:rtp:ms}).
These passes can be freely ordered in the compilation chain without compromising memory safety (\Cref{thm:cccfccdce:rtp:ms}).
The next step in the chain ensures that code stays \gls*{scct} (\Cref{thm:ccscct:rtp:scct}) when compiled from $\Lms$ to $\Lscct$, which is done by switching on a constant-tame mode for the computation.
Lastly, by introducing barriers immediately after branches, speculative leaks via SPECTRE-PHT are prevented when compiling $\Lscct$ to $\Lspec$.
The final result is that the whole compilation chain robustly preserves \gls*{specms} (\Cref{thm:ccall:rtp:specms}).

@@

\subsection{Security Properties and Their Meaning}\label{subsec:formalities:props}
Each of the presented compilers use a cross-language trace relation, which is also used to translate the property from one language to the other one (via the existential or universal images).
% 
%Whenever such translations occur, it is important to check whether there has been a change of meaning.
%Secure compiler engineers need to ensure that such translations of properties do not change meaning in a fundamental wrong way. 
% lifting or lowering reasoning from target- to source-level or vice-versa,
While the meaning of projected properties does change with a translation, the change should not allow for a flawed compilation pipeline.
% 
For example, we could be using a trace relation that translates a property in a language to a totally different one in another language.
% 
To raise the trust into the translation of properties, \Cref{thm:prop-rel-corr} states (in a general fashion) that each security property is faithfully translated using the universal image according to the cross-language trace relation induced by the compiler.

@@

Finally, let us consider a trace $\irl{\trace}\in\mssafe$ and understand what is that is related to via $\sim^{\Ltrg}_{\Lms}$.
% 
{\em All} traces $\trg{\trace}$ with $\trg{\trace}\sim^{\Ltrg}_{\Lms}\irl{\trace}$ are identical to $\irl{\trace}$ except for get and set actions, which require for in-bound accesses (as stated in \Cref{subsec:cs:ms}): this clearly respect $\mssafe$.
%

>>>>

% Fixing: Too many math alphabets used in version normal
\newcommand\hmmax{0}
\newcommand\bmmax{0}
\documentclass[acmsmall]{acmart}
%%
%% \BibTeX command to typeset BibTeX logo in the docs
\AtBeginDocument{%
  \providecommand\BibTeX{{%
    Bib\TeX}}}

%% Rights management information.  This information is sent to you
%% when you complete the rights form.  These commands have SAMPLE
%% values in them; it is your responsibility as an author to replace
%% the commands and values with those provided to you when you
%% complete the rights form.
\setcopyright{acmlicensed}
\copyrightyear{2018}
\acmYear{2018}
\acmDOI{XXXXXXX.XXXXXXX}

%%
%% These commands are for a JOURNAL article.
\acmJournal{JACM}
\acmVolume{37}
\acmNumber{4}
\acmArticle{111}
\acmMonth{8}

%%
%% Submission ID.
%% Use this when submitting an article to a sponsored event. You'll
%% receive a unique submission ID from the organizers
%% of the event, and this ID should be used as the parameter to this command.
%%\acmSubmissionID{123-A56-BU3}

%%
%% For managing citations, it is recommended to use bibliography
%% files in BibTeX format.
%%
%% You can then either use BibTeX with the ACM-Reference-Format style,
%% or BibLaTeX with the acmnumeric or acmauthoryear sytles, that include
%% support for advanced citation of software artefact from the
%% biblatex-software package, also separately available on CTAN.
%%
%% Look at the sample-*-biblatex.tex files for templates showcasing
%% the biblatex styles.
%%

%%
%% The majority of ACM publications use numbered citations and
%% references.  The command \citestyle{authoryear} switches to the
%% "author year" style.
%%
%% If you are preparing content for an event
%% sponsored by ACM SIGGRAPH, you must use the "author year" style of
%% citations and references.
%% Uncommenting
%% the next command will enable that style.
%%\citestyle{acmauthoryear}

@@

\usepackage[colorinlistoftodos]{todonotes}
\usepackage[inference]{semantic}
\usepackage[dvipsnames]{xcolor}
\usepackage[switch]{lineno}
\usepackage{halloweenmath}
\usepackage{fontawesome5}
\usepackage{listofitems}
\usepackage{breakcites}
\usepackage{glossaries}
\usepackage{hyperref}
\usepackage{cleveref}
\usepackage{stmaryrd}
\usepackage{marvosym}
\usepackage{listings}
\usepackage{amsmath}
\usepackage{nameref}
\usepackage{amsthm}
\usepackage{xspace}
\usepackage{xfrac}
\usepackage{tikz}
\usepackage{soul}
\usepackage{bm}
\usepackage{paralist}

@@

\loadglsentries{acronyms}
\makeglossaries
\renewcommand\thelinenumber{\color{red}\arabic{linenumber}}
\begin{document}
\linenumbers
\title{
  Secure Composition 
  \\ 
  of Robust and Optimising Compilers
}

\author{Matthis Kruse}
\affiliation{%
   \institution{Saarland University and CISPA}
   \city{Saarbr\"ucken}
   \state{Saarland}
   \country{Germany}}
\email{matthis.kruse@cispa.de}
\author{Marco Patrignani}
\affiliation{%
   \institution{Trento University}
   \city{Trento}
   \state{South Tyrol}
   \country{Italy}}
\email{marco.patrignani@unitn.it}
%%
%% By default, the full list of authors will be used in the page
%% headers. Often, this list is too long, and will overlap
%% other information printed in the page headers. This command allows
%% the author to define a more concise list
%% of authors' names for this purpose.
\renewcommand{\shortauthors}{Kruse and Patrignani}

@@

%%
%% The code below is generated by the tool at http://dl.acm.org/ccs.cfm.
%% Please copy and paste the code instead of the example below.
%%
\begin{CCSXML}
<ccs2012>
 <concept>
  <concept_id>00000000.0000000.0000000</concept_id>
  <concept_desc>Do Not Use This Code, Generate the Correct Terms for Your Paper</concept_desc>
  <concept_significance>500</concept_significance>
 </concept>
 <concept>
  <concept_id>00000000.00000000.00000000</concept_id>
  <concept_desc>Do Not Use This Code, Generate the Correct Terms for Your Paper</concept_desc>
  <concept_significance>300</concept_significance>
 </concept>
 <concept>
  <concept_id>00000000.00000000.00000000</concept_id>
  <concept_desc>Do Not Use This Code, Generate the Correct Terms for Your Paper</concept_desc>
  <concept_significance>100</concept_significance>
 </concept>
 <concept>
  <concept_id>00000000.00000000.00000000</concept_id>
  <concept_desc>Do Not Use This Code, Generate the Correct Terms for Your Paper</concept_desc>
  <concept_significance>100</concept_significance>
 </concept>
</ccs2012>
\end{CCSXML}

\ccsdesc[500]{Do Not Use This Code~Generate the Correct Terms for Your Paper}
\ccsdesc[300]{Do Not Use This Code~Generate the Correct Terms for Your Paper}
\ccsdesc{Do Not Use This Code~Generate the Correct Terms for Your Paper}
\ccsdesc[100]{Do Not Use This Code~Generate the Correct Terms for Your Paper}

%%
%% Keywords. The author(s) should pick words that accurately describe
%% the work being presented. Separate the keywords with commas.
\keywords{Do, Not, Us, This, Code, Put, the, Correct, Terms, for,
  Your, Paper}

\received{20 February 2007}
\received[revised]{12 March 2009}
\received[accepted]{5 June 2009}

@@

  \item %
  		This paper provides formal proofs of the security properties preserved by each of the presented passes. 
  		% 
  		Additionally this paper analyses each pass in terms of their compatibility with each other (\Cref{sec:formalities}), proving how each pass fulfills the conditions that make their composition meaningful from a security standpoint.
        % In doing so, the paper provides formal insights into some technicalities of the concrete proofs necessary as technical setup to reap the benefits of the presented compositionality framework.

@@

Properties $\pi$ are sets of traces of admissible program behaviours, ascribing what said property considers valid.
% 
The set of all properties can be split-up into different {\em classes} ($\class$), i.e., safety, liveness, and neither safety nor liveness~\cite{clarkson2008hyper}, so a class is simply a set of properties.
% 
The compositionality framework (\Cref{sec:sequential}) presented in this paper considers arbitrary classes, while the case-study (\Cref{sec:casestud:rtp}) fixes them to concrete instances of safety properties, since it is decidable whether a trace satisfies a safety property with just a finite trace (i.e., a \emph{prefix}).
As an example, consider the trace:
$$\ev{Dealloc\ \loc}\cdot\ev{Read\ \loc\ 1729}\cdot\dots$$
which describes the interaction with a memory where the deallocation of an address $\loc$ precedes a read (of some value $1729$) at that address in memory.
% 
This program behaviour is insecure w.r.t. a canonical notion of (temporal) memory safety dictating no use-after-frees of pointers~\cite{nagarakatte2010cets,azevedo2018meaningsofms}, because it reads from a memory location that was freed already.
The previous finite trace prefix is enough to decide that the trace does not satisfy \gls*{tms} and there is no way to append events to this prefix which would result in the trace being admissible.

@@

This paper uses a general notion of robust preservation~\cite{abate2021extacc} that considers compilers that use languages with different trace models. 
To this end, considering a source trace $\src{\trace}$ and a target trace $\trg{\trace}$, there is a relation ($\sim$) describing the effect of a corresponding compiler (see \Cref{sec:casestud:rtp}). 
This relation induces the following two projection functions~\cite{abate2021extacc}: (1) the \emph{existential image} $\tau_\sim\left(\src{\pi}\right)$ and (2) the \emph{universal image} $\sigma_\sim\left(\trg{\pi}\right)$.
These projections map source-level (resp. target-level) properties to target-level (resp. source-level) properties in a way that identifies the ``same'' property across languages. 
The case study of this paper uses the universal image, since some considered properties, such as \gls*{ss}, are not definable in a higher-level language that, e.g., does not model speculation.
% For self-containment, the universal image is defined as follows~\cite{abate2021extacc}:
% 
\begin{definition}[Universal Image]
% [Existential and Universal Image]
\label{def:universal:img}\label{def:sigma}
  % \[
  %   \tau_\sim\left(\src{\pi}\right) := 
  %     \left\{ 
  %       \trg{\trace} \mid \exists \src{\trace}\ldotp \src{\trace}\sim\trg{\trace}, \text{ and } \src{\trace}\in\src{\pi} 
  %     \right\}
  % \]
  \[ 
    \sigma_\sim\left(\trg{\pi}\right) := 
      \left\{ 
        \src{\trace} \mid \forall \trg{\trace}\ldotp \text{if }\src{\trace}\sim\trg{\trace}, \text{ then } \trg{\trace}\in\trg{\pi} 
      \right\}
  \]
\end{definition}
\noindent
With this projection function, we define a more general version of robust preservation as follows~\cite{abate2021extacc}.
% For a \bul{compiler $\cc{\src{L}}{\trg{L}}$ to robustly preserve a class of source properties $\src{\class}$}, if for any \rul{property $\src{\pi}$ of that class $\src{\class}$ and programs $\src{p}$ written in $\src{L}$} where \iul{$\src{\progvar}$ robustly satisfies $\src{\pi}$}, then \oul{the compilation of $\src{\progvar}$, $\cc{\src{L}}{\trg{L}}\left(\src{p}\right)$, must robustly satisfy $\tau_\sim\left(\src{\pi}\right)$}.
% 
% \begin{definition}[Robust Preservation with $\tau_\sim$]\label{def:rtp:tau}
%   % $\;$\\
%   % \vspace{-1em}
%   % \begin{nscenter}
%   % Compiler $\cc{\src{L}}{\trg{L}}$ robustly preserves $\class$, 
%   \bul{$\rtptau{\cc{\src{L}}{\trg{L}}}{\src{\class}}{\sim}$}
%   %, iff 
%   $\isdef$
%   \rul{$\forall \src{\pi}\in\src{\class}, \src{p}\in\src{L},$} if \iul{$\rsat{\src{\progvar}}{\src{\pi}}$}, then \oul{$\rsat{\cc{\src{L}}{\trg{L}}\left(\src{p}\right)}{\tau_\sim{\left(\src{\pi}\right)}}$}.
%   % \end{nscenter}
% \end{definition}
% 
A \bul{compiler $\cc{\src{L}}{\trg{L}}$ robustly preserves a class of target properties $\trg{\class}$}, if for any \rul{property $\trg{\pi}$ of class $\trg{\class}$ and programs $\src{p}$}, where \iul{$\src{\progvar}$ robustly satisfies $\sigma_\sim\left(\trg{\pi}\right)$}, \oul{for the compilation of $\src{\progvar}$, we have that $\cc{\src{L}}{\trg{L}}\left(\src{p}\right)$ robustly satisfies $\trg{\pi}$}.

@@

Let \bul{$\cc{\src{L}}{\trg{L}}$ robustly preserve the class $\sigma_{\sim_2}\left(\obj{\class[_1]}\right)$ under $\sim_1$} and let \rul{$\cc{\trg{L}}{\obj{L}}$ robustly preserve the class $\obj{\class[_2]}$ under $\sim_2$}.
Then, \iul{when the cross-language relation $\sim_2$ is well-formed w.r.t. class $\obj{\class[_1]}$}, it follows that \oul{the composed compiler\footnote{Here, $\cc{\src{L}}{\trg{L}}\circ\cc{\trg{L}}{\obj{L}}$ means first running $\cc{\src{L}}{\trg{L}}$ and then $\cc{\trg{L}}{\obj{L}}$.} $\cc{\src{L}}{\trg{L}}\circ\cc{\trg{L}}{\obj{L}}$ robustly preserves the intersection of classes $\obj{\class[_1]}\cap\obj{\class[_2]}$ under $\sim_1\bullet\sim_2$}.
% 
%For space reasons, \Thmref{thm:urtp} and \Thmref{thm:lrtp} are not presented here, but can be found in the technical appendix ($\CoqSymbol$).
% Note that it is not necessary to require $\sim_2$ to be well-formed w.r.t. $\src{\class[_1]}$, since the 
% \MPin{why}
\begin{theorem}[Composition of Secure Compilers w.r.t. $\sigma$]\label{thm:rtpsim:sig}
  $\;$

  If \bul{$\rtpsig{\cc{\src{L}}{\trg{L}}}{\tilde{\sigma}_{\sim_2}\left(\obj{\class[_{1}]}\right)}{\sim_1}$}, \rul{$\rtpsig{\cc{\trg{L}}{\obj{L}}}{\obj{\class[_2]}}{\sim_2}$}, and \iul{$\wfcsig{\sim_2}{\obj{\class[_1]}}$}, \\ then \oul{$\rtpsig{\cc{\src{L}}{\trg{L}}\circ\cc{\trg{L}}{\obj{L}}}{\obj{\class[_{1}]}\cap\obj{\class[_{2}]}}{\sim_1\bullet\sim_2}$}. \Coqed
\end{theorem}

@@

This section introduces trace properties of interest for this paper: \gls*{tms}, \gls*{sms}, \gls*{ms}, \gls*{scct}, and \gls*{ss}.
These properties are of practical importance (as mentioned in \Cref{sec:introduction}) and also of interest in the case study presented later (\Cref{sec:casestud:defs,sec:casestud:rtp}). 
This section presents all of them, despite the fact that they are inspired by existing work, in order to showcase all that is required for a formal proof of security for a realistic compilation toolchain.
%\MP{ say something about this in the rebuttal: only the defs are borrowed, the universal images and all that work is new}
%\MK{ universal images are not new. }
% , none of the properties is novel, for self-containment of the paper, their definition is necessary for the case study.
% For each of the presented properties, the technical report defines corresponding monitors, which refine a given property and have a key role in the proofs of this paper. 
The technical report defines monitors for each of the presented properties.
% 
Monitors refine each property and have a key role in the proofs of this paper.
% managed to prove 
%  if tms-monitor --[trace]-->* ∅  
%  then trace ∈ tms
% but not
%  if trace ∈ tms
%  then tms-monitor --[trace]-->* ∅

@@

To express Constant Time, we extend the memory safety trace model with a {\em security tag} ($\securitytag{}$) that indicates whether events contain sensitive information ($\lock$) or not ($\unlock$).

@@

\[
  \typerulenolabel{specrel:drop:tag}{}{\preevent[_{\ctsafe}];\lock\sim_{\mathghost}\preevent[_{\mathghost}];\lock_{\text{NONE}}}
  \typerulenolabel{specrel:drop:spectag}{}{\emptyevent\sim_{\mathghost}\preevent[_{\mathghost}];\lock_{\text{PHT}}}
\]
\[
  \typerulenolabel{specrel:drop:pub}{}{\preevent[_{\ctsafe}];\unlock\sim_{\mathghost}\preevent[_{\mathghost}];\unlock}
  \typerulenolabel{specrel:drop:crash}{}{\lightning\sim_{\mathghost}\lightning}
\]
\[
  \typerulenolabel{specrel:drop:spec}{}{\emptyevent\sim_{\mathghost}\ev{Spec}}
  \typerulenolabel{specrel:drop:rlb}{}{\emptyevent\sim_{\mathghost}\ev{Rlb}}
  \typerulenolabel{specrel:drop:barrier}{}{\emptyevent\sim_{\mathghost}\ev{Barrier}}
\]

@@

% \vspace{-1em}
\[
  \begin{array}{rrcl}
    \text{(Values)} & \trg{v} &:=& \dots \mid \trg{\lpair{v_1}{v_2}} \\
    \text{(Expressions)} & \trg{\expr} &:=& \dots \mid \trg{\lhast{\expr}{\tau}} \mid \trg{\lispoisoned{x}} \mid \trg{e.1} \mid \trg{e.2}
  \end{array}
\]

@@

    \draw[dashed] ($(S)-(0.0,-0.5)$) -- ($(S)-(0.0,-2.5)$);
    \draw[dashed] ($(T)-(0.0,-0.5)$) -- ($(T)-(0.0,-2.5)$);
    \draw[dashed] ($(M)-(0.0,-0.5)$) -- ($(M)-(0.0,-2.5)$);
    \draw[dashed] ($(E)-(0.0,-0.5)$) -- ($(E)-(0.0,-2.5)$);
    \draw[dashed] ($(O)-(0.0,-0.5)$) -- ($(O)-(0.0,-2.5)$);
    \draw[dashed] ($(Z)-(0.0,-0.5)$) -- ($(Z)-(0.0,-2.5)$);
  \end{tikzpicture}
  \vspace{-1em}
  \caption{Visualisation of the optimising compilation pipeline that preserves \gls*{specms}. %
    Vertices in the graph are the programming languages from earlier sections (\Cref{sec:casestud:defs}). %
    Full edges are secure compilers passes.
    Dotted edges are composition of passes and use the presented framework (\Cref{sec:sequential}) to indicate the property they preserve. %
    The dashed lines partition the graph into the sections where the respective theorems are presented.
  }\label{fig:pipeline}
\end{figure*}
The section demonstrates the power of the framework (\Cref{sec:sequential}) by composing these compilers for a secure and optimising compilation chain that robustly preserves \gls*{specms}.
The first step in this chain is the compiler from $\src{L_{\tmssafe}}$ to $\trg{L}$ that robustly preserves just \gls*{tms} (\Cref{thm:cca:rtp:tms}).
From here, another pass from $\trg{L}$ to $\irl{L_{\mssafe}}$ ensures that no out-of-bounds accesses can happen and, thus, programs at this point attain \gls*{sms} (\Cref{thm:ccb:rtp:sms}).
Since these properties compose into \gls*{ms}, composing these passes yields a compiler that robustly preserves \gls*{ms} (\Cref{corr:ccab:rtp:ms}).
Then, the section presents two optimisation passes, namely \gls*{cf} and \gls*{dce}, each of which robustly preserves \gls*{ms} (\Cref{thm:ccdce:rtp:ms,thm:cccf:rtp:ms}).
These passes can be freely ordered in the compilation chain without compromising memory safety (\Cref{thm:cccfccdce:rtp:ms}).
The next step in the chain ensures that code stays \gls*{scct} (\Cref{thm:ccscct:rtp:scct}) when compiled from $\Lms$ to $\Lscct$, which is done by switching on a constant-time mode for the computation.
Lastly, by introducing barriers immediately after branches, speculative leaks via SPECTRE-PHT are prevented when compiling $\Lscct$ to $\Lspec$.
The final result is that the whole compilation chain robustly preserves \gls*{specms} (\Cref{thm:ccall:rtp:specms}).

@@

\subsection{Security Properties and Their Meaning}\label{subsec:formalities:props}
Each of the presented compilers uses a cross-language trace relation, which is also used to translate the property from one language to the other one (via the existential or universal images).
% 
%Whenever such translations occur, it is important to check whether there has been a change of meaning.
%Secure compiler engineers need to ensure that such translations of properties do not change meaning in a fundamental wrong way. 
% lifting or lowering reasoning from target- to source-level or vice-versa,
While the meaning of projected properties does change with a translation, the change should not allow for a flawed compilation pipeline.
% 
For example, we could be using a trace relation that translates a property in a language to a totally different one in another language.
% 
To raise the trust into the translation of properties, \Cref{thm:prop-rel-corr} states (in a general fashion) that each security property is faithfully translated using the universal image according to the cross-language trace relation induced by the compiler.

@@

Finally, let us consider a trace $\irl{\trace}\in\mssafe$ and understand what is related to via $\sim^{\Ltrg}_{\Lms}$.
% 
{\em All} traces $\trg{\trace}$ with $\trg{\trace}\sim^{\Ltrg}_{\Lms}\irl{\trace}$ are identical to $\irl{\trace}$ except for get and set actions, which require for in-bound accesses (as stated in \Cref{subsec:cs:ms}): this clearly respect $\mssafe$.
%
